Supporting Information for\\ ``A weakly structured stem for human origins in Africa''

### 2022-07-22 supplement.tex
fix OOA parameters

In our preliminary models, the unconstrained inferred split times between East
African populations and Europeans was inferred to have occurred around $40$ka,
which is more recent than we may have expected given the well-documented
presence of modern humans outside of Africa at that time \citep{}, as well as
previous genetics-based dating of splits between African and Eurasian
populations \citep{}. The unconstrained inferred split between East and West
African populations occurred about $10$ka earlier
(Section~\ref{sec:supp-results}).
%We see two non-exclusive explanations for this apparent discrepancy. First,
%there is uncertainty in our model fit of this parameter. The $95\%$ confidence
%interval allows for a split as far back as \sgcomment{$XX$}.
Across all models, the $95\%$ confidence intervals for the OOA event was
roughly $38 \pm 5$ka, and for the East/West African split was $45 \pm 10$ka.
However, our confidence intervals do not capture all sources of uncertainty.
For example, model mis-specification can lead to a mismatch between an inferred
split time and the first time populations actually split. Our model supposes a
single OOA event followed by constant symmetric migrations between populations,
which is certainly an oversimplification \citep{}. Additionally, most previous
estimates of the time of the OOA expansion compared Eurasian and West African
populations. Our inferred time of the East/West African divergence aligns with
previous estimates of the OOA event at $\approx 60$ka (themselves inferred from
oversimplified demographic models). Rather than adding the many parameters
required to account for these effects, we fixed the split time between
Europeans and the East African ancestral population at $50$ka, and the split
between East/West African population at $60$ka.
%We discuss the sensitivity of our model to these assumptions in Section
%\ref{sec:change-dates}.

>>>>

In our preliminary models, the unconstrained inferred split times between East
African populations and Europeans was inferred to have occurred around $40$ka,
which is more recent than we may have expected given the well-documented
presence of modern humans outside of Africa at that time \citep{}, as well as
previous genetics-based dating of splits between African and Eurasian
populations \citep{}. The unconstrained inferred split between East and West
African populations occurred about $10$ka earlier
These recent split dates seems to be at odds with archaelogical evidence for earlier
 Out-of-Africa (OOA) expansion, and with the estimated date for Neanderthal admixture
 into OOA populations.  The parameter uncertainty due to finite genome
 do not account for this difference:  in the high-likelihood models, the $95\%$ confidence 
 intervals for the east African-European split was roughly $38 \pm 5$ka, 
 and for the East/West African split was $45 \pm 10$ka, with the uncertainties 
 reflecting 2 standard deviations according to bootstrap of genomic regions.  
Thus we expect these discrepancies to result for model misspecification in the 
details of the divergence between the ancestors of OOA and east African 
populations. Our models supposes a
single split prior to the OOA event followed by constant symmetric migrations 
between eastern and OOA populations.
which is certainly an oversimplification \citep{}.

Since a detailed model for the Out of Africa and Back-to-Africa was not our main focus, 
we chose to keep this simplified model, but fix the split between east African-European 
split at $50ka.$ This allowed us to include Neanderthal admixture into the OOA 
population at the previously-estimated $45ka$. Fixing this parameter pushed back the 
estimated value for the split between the ancestors of east and West African populations at around 
$60ka$, a value that we took as fixed in most subsequent analyses.

We discuss the sensitivity of our model to these assumptions in Section
\ref{sec:change-dates}.

\sgcomment{I feel like we should improve the discussion of relationship with previous estimates below, or remove.}

In addition to archaeological evidence, previous genetic dating suggested 
earlier dates for a split between European and African populations. Most of 
these results focused on populations of recent West African ancestry.

Additionally, most previous
estimates of the time of the OOA expansion compared Eurasian and West African
populations. Our inferred time of the East/West African divergence aligns with
previous estimates of the OOA event at $\approx 60$ka (themselves inferred from
oversimplified demographic models).

%We discuss the sensitivity of our model to these assumptions in Section
%\ref{sec:change-dates}.
